---
title: "Introduction to `ggplot2` homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

# Homework Info

This homework recaps what you learnt in class today. 

You'll be looking at the `backpack` dataset contained within the `CodeClanData` package using tidyverse. You'll need to load the appropriate packages first.In all plots you make, ensure your labels are appropriate 

**You must write down a short summary of what each plot tells you in order to meet MVP.** Being an analyst is not just the ability to write code, you need to be able to interpret the visualisations and translate the relationships found (or not found) into plain language to explain to non-technical decision makers. Getting as much practice of this as you can while on this course is just (if not more!) important than the code!



# MVP

<br>

**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions.   

```{r}
library(tidyverse)
library(ggplot2)
library(CodeClanData)

backpack <- CodeClanData::backpack

backpack %>% 
  names()

backpack %>% 
  dim()
```


**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution. 

```{r}
ggplot(backpack) +
  aes(backpack_weight) +
  geom_bar() +
  labs(x = "backpack weight (kg)",
       y = "number of students") +
  scale_y_continuous(breaks = seq(0, 12, 1)) +
  scale_x_continuous(breaks = seq(0, 36, 5)) +
  coord_flip() +
  theme_minimal()
```
- It is clear that most students have backpacks between 5 and 15 kg, which is
something to consider when analysing the following graphs


**Question 3.**  

Are there differences in backpack weight between sex?   

```{r}
ggplot(backpack) +
  aes(backpack_weight, fill = sex) +
  geom_bar(color = "black", position = position_dodge(preserve = "single")) +
  labs(x = "\nbackpack weight (kg)",
       y = "number of students\n") +
  scale_y_continuous(breaks = seq(0, 10, 1)) +
  scale_x_binned(n.breaks = 18, limits = c(0, 36)) +
  theme_minimal()
```
- We see that males tend to have heavier bags than females (up until 10kg
there are more or an equal amount of females than males with that weight of bag)


**Question 4.**  

What is the relationship between body weight and backpack weight?

```{r}
ggplot(backpack) +
  aes(x = body_weight, y = backpack_weight) +
  geom_point() +
  geom_smooth() +
  theme_minimal()
```
- There is a clear slight positive trend between backpack weight and body weight,
however after about 200lbs this stops correlating as strongly (probably because
body weights aren't commonly heavier than that so we have less data)


**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  

```{r}
ggplot(backpack) +
  aes(x = body_weight, y = backpack_weight, color = sex) +
  geom_point() +
  geom_smooth() +
  facet_grid(~sex)
  theme_minimal()
```
- Not particularly, there is still the positive trend for both genders
- For men, the trend positive trend lasts longer (up until about 200lbs) just
because weights are, in general, higher


**Question 6.**  

What **proportion** of students report back pain?   

```{r}
backpack %>% 
  summarise(back_problems_percentage = 100*mean(back_problems))
```


**Question 7.**  

Is this proportion different between men and women? 

```{r}
backpack %>%
  group_by(sex) %>% 
  summarise(back_problems_percentage = 100*mean(back_problems))
```


**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   

*Hint: faceting would be useful here*.   

```{r}
ggplot(backpack) +
  aes(x = backpack_weight, y = back_problems, fill = sex) +
  geom_col() +
  geom_smooth(se=F, colour = "grey50") +
  facet_grid(~sex) +
  scale_x_continuous(limits = c(0, 26)) +
  theme_minimal()


ggplot(backpack) +
  aes(x = body_weight, y = back_problems) +
  geom_col(aes(fill = sex)) +
  geom_smooth(se=F, colour = "grey50") +
  facet_grid(~sex) +
  scale_x_binned(n.breaks = 15, limits = c(100, 220)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, size = 7)) +
  labs(title = "back problems vs. body weight for different genders\n",
       x = "\nbody weight (lbs)",
       y = "number of students with back problems\n")


ggplot(backpack) +
  aes(x = ratio, y = back_problems, colour = sex) +
  geom_smooth() +
  facet_grid(~sex) +
  theme_minimal()
```
Graph 1:
- We see that there are more females getting back problems around 6-16kg,
however the smoothed line shows that this is due to there being more data for 
that weight of bag and that there is consistently more back problems at 16kg+
bags

Graph 2:
- For males there is a fairly clear peak in back problems around 160/170lbs
body weight
- For females it seems we have two peaks: at about 120lbs and 160/170lbs

Graph 3:
- For females there is a clear positive correlation between the ratio of
bag/body weight and back problems
- For males there is a much less strong relationship, however this could be
because there are not many males with back problems


# Extensions

<br>

**Question 1.** 

Does reported backpain differ across the year groups? Ignore year 0 in your data.  

```{r}
backpack %>% 
  mutate(year = factor(year, levels = c(0,1,2,3,4,5,6))) %>% 
  ggplot() +
  aes(x = year, y = back_problems, fill = sex) +
  geom_col() +
  scale_x_discrete(breaks = seq(1, 6)) +
  theme_minimal() +
  labs(title = "back problems vs. year of study\n",
       y = "no. of back problems\n",
       x = "\nyear")
```


**Question 2.** 

Is there a difference between how heavy the bags those in different years carry?

```{r}

```

