---
title: "Homework - features and elements of multiple regression"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

<hr>

# MVP

<br>

1. Load the `housing_prices.csv` data set and undertake an initial exploration of the data. You will find details on the data set on the relevant [Kaggle page](https://www.kaggle.com/camnugent/california-housing-prices)

```{r}
library(GGally)
library(tidyverse)

housing <- read_csv("data/housing_prices.csv") %>% 
  janitor::clean_names()

head(housing)
```

<br>

2. We expect the `total_rooms` of houses to be strongly correlated with `total_bedrooms`. Use `ggpairs()` to investigate correlations between these two variables.

```{r}
ggpairs(housing[,c("total_rooms", "total_bedrooms")])
```
It has a very strong correlation coeff of 0.934

<br> 

3. So, we do find significant correlations. Let's drop `total_bedrooms` from the dataset, and use only `total_rooms` going forward.

```{r}
housing_tidy <- housing %>% 
  select(-total_bedrooms)
```

<br>

4. We are interested in developing a regression model for the `median_house_value` of a house in terms of the possible predictor variables in the dataset. 

  i. Use `ggpairs()` to investigate correlations between `median_house_value` and the predictors (this may take a while to run, don't worry, make coffee or something).

```{r, message = FALSE}
ggpairs(housing_tidy)
```

  ii. Perform further `ggplot` visualisations of any significant correlations you find.

```{r}
housing_tidy %>% 
  ggplot() +
  aes(x = median_income, y = median_house_value) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```

```{r}
housing_tidy %>% 
  ggplot() +
  aes(x = latitude, y = median_house_value) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```

```{r}
housing_tidy %>% 
  ggplot() +
  aes(x = total_rooms, y = median_house_value) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```

```{r}
housing_tidy %>% 
  ggplot() +
  aes(x = ocean_proximity, y = median_house_value) +
  geom_boxplot()
```

<br>

5. Shortly we may try a regression model to fit the categorical predictor `ocean_proximity`. Investigate the level of `ocean_proximity` predictors. How many dummy variables do you expect to get from it?  
  
There are 5 levels so we'd expect to get 4 dummy variable as one would be inferred by the other 4.

<br>
 
6. Start with simple linear regression. Regress `median_house_value` on `median_income` and check the regression diagnostics.

```{r}
model1 <- lm(median_house_value ~ median_income, data = housing_tidy)
plot(model1)
```

<br> 

7. Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.

```{r}
model2 <- lm(median_house_value ~ median_income + total_rooms, data = housing_tidy)

plot(model2)
```

# Extension
    
<br>

8. Try adding an interaction between `log(medium_income)` and your chosen categorical predictor. Do you think this interaction term is statistically justified?

```{r}
model3 <- lm(log(median_house_value) ~ log(median_income) + ocean_proximity + log(median_income):ocean_proximity, data = housing_tidy)
summary(model3)

anova(model2, model3)
```

<br>

9. Find and plot an appropriate visualisation to show the effect of this interaction

```{r}

```


